% SPDX-FileCopyrightText: Copyright (c) 2020-2026 Yegor Bugayenko
% SPDX-License-Identifier: MIT

\documentclass[acmsmall,nonacm,11pt]{acmart}
\settopmatter{printfolios=false,printccs=false,printacmref=false}
\makeatletter\let\@authorsaddresses\@empty\makeatother
\usepackage[noframes,nonumbers,nobars,samepage]{ffcode}
% listings emits \vspace around each display, which acmart warns about;
% restore the original \vspace while a listing is typeset
\makeatletter
\lst@AddToHook{DisplayStyle}{\let\@vspace\@vspace@orig\let\@vspacer\@vspacer@orig}
\makeatother
\usepackage{natbib}
\usepackage{doi}
\usepackage{multicol}
\usepackage{eolang}
\usepackage{href-ul}
\normalem
\usepackage[capitalize]{cleveref}
\usepackage{to-be-determined}
\usepackage[utf8]{inputenc}
\usepackage[T2A,T1]{fontenc}
\usepackage[russian,english]{babel}
  \DeclareFontFamilySubstitution{T2A}{LinuxBiolinumT-TLF}{LibertinusSans-TLF}
  \renewcommand\ttdefault{cmtt}
% stmaryrd (loaded by eolang) lacks a 5.5pt design size for scriptscript math
\DeclareFontFamily{U}{stmry}{}
\DeclareFontShape{U}{stmry}{m}{n}{%
  <5><6><7><8><9><10>gen*stmary%
  <5.5>stmary5%
  <10.95><12><14.4><17.28><20.74><24.88>stmary10%
}{}
\usepackage{csquotes}
% see https://tex.stackexchange.com/a/256190/1449
\usepackage{url}
  \def\UrlBreaks{\do\/\do-}
\setlength{\footskip}{13.0pt}

\newsavebox\ffboxa
\newsavebox\ffboxb
\newdimen\ffmaxd
\newcommand\ffsep{%
  \hspace{0.4em}\textcolor{gray}{\vrule width 0.3pt depth \ffmaxd}\hspace{0.4em}}
\newcommand\twoffinputs[3]{%
  \sbox\ffboxa{\parbox[t]{#1\linewidth}{\vspace{0pt}\ffinput{#2}}}%
  \sbox\ffboxb{\parbox[t]{\linewidth - #1\linewidth - 1em}{\vspace{0pt}\ffinput{#3}}}%
  \ffmaxd=\dp\ffboxa
  \ifdim\dp\ffboxb>\ffmaxd \ffmaxd=\dp\ffboxb \fi
  \noindent\usebox\ffboxa\ffsep\usebox\ffboxb}

\newcommand\deff[1]{\ff{#1}}
\newcommand\adeff[1]{\ff{#1}}

\newcommand\broken{\tbd{This section is not valid, must be rewritten!}}

\title{Reducing Programs to Objects}
\subtitle{
  Ver:
  \texorpdfstring{
    \href{https://github.com/REPOSITORY/releases/tag/0.0.0}
      {\ff{0.0.0}}
  }{0.0.0}
}
\author{Yegor Bugayenko}
\affiliation{\institution{Huawei}\country{Russia}\city{Moscow}}
\email{yegor256@gmail.com}

\begin{abstract}
C++, Java, C\#, Python, Ruby, and JavaScript are the most
  powerful object-oriented programming languages if language power
  is defined as the number of \emph{features} available to a programmer.
\eolang{}, on the other hand, is an object-oriented programming language
  with a reduced set of features: it has nothing but objects
  and mechanisms of their composition and decoration.
This paper asks the following research question: ``Which known features
  are possible to implement using only objects?''
\end{abstract}

\begin{document}
\raggedbottom

\maketitle

\section{Introduction}

We selected the most complex features and demonstrated how each
  of them may be represented in \eolang{}~\citep{bugayenko2021eolang}
  objects%
    \footnote{%
    \LaTeX{} sources of this paper are maintained in
    \href{https://github.com/REPOSITORY}{REPOSITORY} GitHub repository,
  the rendered version is
    \href{https://github.com/REPOSITORY/releases/tag/0.0.0}{\ff{0.0.0}}.}.

Other features are more trivial, which is why they are not presented
  in this paper, such as operators, loops, variables, code blocks,
  constants, and selection.

The work of \citet{bugayenko2022origins} explains the design of \eolang{}
  standard library of objects, which are used below.

\section{Goto}
\label{sec:goto}

A ``backward'' jump in C language can be mapped to \eolang{} recursion:

\begin{ffsave}[goto-backward.c]
#include "stdio.h"
void f() {
  int i = 1;
  again:
  i++;
  if (i < 10)
    goto again;
  printf("Finished!");
}
\end{ffsave}
\begin{ffsave}[goto-backward.eo]
malloc.empty > [] > f
  seq * > [i]
    i.put 1
    again i
    io.stdout "Finished!"
  seq * > [ii] > again
    ii.put (ii.as-number.plus 1)
    if.
      ii.as-number.lt 10
      again ii
      true
\end{ffsave}
\twoffinputs{0.4}{goto-backward.c}{goto-backward.eo}

A ``forward'' jump in C language can be mapped to \eolang{}:

\begin{ffsave}[goto-forward.c]
int f(int x) {
  int r = 0;
  if (x == 0)
    goto exit;
  r = 42 / x;
  exit:
  return r;
}
\end{ffsave}
\begin{ffsave}[goto-forward.eo]
malloc.empty > [x] > f
  seq * > [r] >>
    r.put 0
    if.
      x.eq 0
      true
      r.put (42.div x)
    number r.get
\end{ffsave}
\twoffinputs{0.4}{goto-forward.c}{goto-forward.eo}

According to the B\"{o}hm–Jacopini theorem~\citep{bohm1966flow},
  any use of the \ff{GOTO} operator may be replaced
  by a combination of
  sequence (the \deff{seq} object),
  selection (the \deff{if} attribute),
  and
  iteration (the \deff{while} object).
Relying on this fact, we may conclude that any C code that uses \ff{GOTO}
  may be represented in \eolang{}.

\section{Pointers}
\label{sec:pointers}

This is an example of C code, where the pointer \ff{p} is first incremented
  by seven times \ff{sizeof(book)} (which is equal to 108)
  and then de-referenced to become a struct \ff{book} mapped to memory.
Then, the \ff{title} part of the struct is filled with a string
  and the \ff{price} part is returned as a result of the function \ff{f}.
The algorithm can be represented in \eolang{} on top of the \adeff{malloc}.

\begin{ffsave}[pointer-data.c]
#include "string.h"
struct book {
  char title[100];
  int price;
};
int f(struct book* p) {
  struct book b = *(p + 7);
  strcpy(b.title, "1984");
  return b.price;
}
\end{ffsave}
\begin{ffsave}[pointer-data.eo]
[mem offset] > book
  book > [i] > plus
    mem
    offset.plus (i.times 108)
  as-i32. > get-price
    mem.read (offset.plus 100) 4
  mem.write offset s > [s] > set-title
[p] > f
  (book p).plus 7 > b
  seq * > @
    b.set-title "1984"
    b.price
\end{ffsave}
\twoffinputs{0.4}{pointer-data.c}{pointer-data.eo}

A pointer may not only refer to data in the heap but also
  to executable code in memory (there is no difference between
  ``program'' and ``data'' memory in both x86 and ARM architectures).
This is an example of a C program that calls a function referenced
  by a function pointer, providing two integer arguments to it.
It can easily be mapped to \eolang{}, via its partial application:

\begin{ffsave}[pointer-fn.c]
typedef int foo(int, int);
int f(foo *p) {
  return (*p)(7, 42);
}
\end{ffsave}
\begin{ffsave}[pointer-fn.eo]
[x y] > foo

[foo] > f
  foo 7 42
\end{ffsave}
\twoffinputs{0.4}{pointer-fn.c}{pointer-fn.eo}

It is important to note that the following code
  cannot be represented in \eolang{}:

\begin{ffcode}
int f() {
  int (*p)(int);
  p = (int(*)(int)) 0x761AFE65;
  return (*p) (42);
}
\end{ffcode}

Here, the pointer refers to an arbitrary address in program memory,
  where some executable code is supposed to be located.

This C function, which returns seven (tested with GCC),
  assumes that variables are placed in the stack sequentially
  and uses pointer de-referencing to access them.

\begin{ffsave}[pointer-stack.c]
long f() {
  long a = 42;
  long b = 7;
  return *(&a - 1);
}
\end{ffsave}
\begin{ffsave}[pointer-stack.eo]
malloc.empty > [] > f
  seq * > [m]
    m.write 0 7
    m.write 8 42
    number
      m.read (8.minus 8) 8
\end{ffsave}
\twoffinputs{0.4}{pointer-stack.c}{pointer-stack.eo}

\section{Impure Functions}

A function is pure when, among other qualities,
  it does not have side effects: no mutation of local static variables,
  non-local variables, mutable reference arguments, or input/output streams.
This impure PHP function can be mapped to \eolang{}:

\begin{ffsave}[impure.php]
$a = 42;
function inc() {
  $a = $a + 1;
  return $a;
}
inc(inc($a)); // $a == 44
\end{ffsave}
\begin{ffsave}[impure.eo]
[a] > inc
  number a.get > ia
  seq * > @
    a.put (ia.plus 1)
    ia
\end{ffsave}
\twoffinputs{0.5}{impure.php}{impure.eo}

Here, \ff{a} is an abstract of a memory block.
It must be passed as an argument to the \ff{inc} object.
In PHP, and many other languages, the heap is available in global state,
  while in \eolang{} it must be passed explicitly as an argument.

\section{Classes}
\label{sec:classes}

\broken

A class is an extensible program code template for creating objects, providing initial values for state (member variables) and implementations of behavior (member functions or methods).
The following program contains a Ruby class with a single constructor, two methods, and one mutable member variable:

\begin{ffcode}
class Book
  def initialize(i)
    @id = id
    puts "New book!"
  end
  def path
    "/tmp/${@id}.txt"
  end
  def move(i)
    @id = i
  end
end
\end{ffcode}

It can be mapped to the following \eolang{} code, where a class becomes an object factory:

\begin{ffcode}
[i] > book
  cage i > id
  [] > ruby-init
    tt.sprintf > @
      "New book!"
  [] > path
    tt.sprintf > @
      "/tmp/%s.txt"
      id
  [i] > move
    id.write i > @
\end{ffcode}

Here, the constructor is represented by the object \ff{ruby-init}, which finishes the initialization of the object.
Making an ``instance'' of a book would look like this:

\begin{ffcode}
book 42 > b
b.ruby-init
b.move 7
b.path
\end{ffcode}

\section{Destructors}
\label{sec:destructors}

\broken

In C++ and some other languages, a destructor is a method that is called for a class object when that object passes out of scope or is explicitly deleted.
The following code will print both \ff{Alive} and \ff{Dead} texts:

\begin{ffcode}
#include <iostream>
class Foo {
public:
  Foo() { std::cout << "Alive"; }
  ~Foo() { std::cout << "Dead"; };
};
int main() {
  Foo f = Foo();
}
\end{ffcode}

It may be translated to \eolang{} as such:

\begin{ffcode}
[] > foo
  [] > constructor
    io.stdout "Alive" > @
  [] > destructor
    io.stdout "Dead" > @
[] > main
  foo > f
  seq * > @
    f.constructor
    f.destructor
\end{ffcode}

There is no garbage collection in \eolang{}, so a destructor must be explicitly ``called'' when an object passes out of scope or is deleted.

\section{Exceptions}
\label{sec:exceptions}

Exception handling is the process of responding to the occurrence of exceptions—anomalous or exceptional conditions requiring special processing—during the execution of a program.
This C++ program utilizes exception handling to avoid segmentation fault due to null pointer de-referencing:

\begin{ffcode}
#include <iostream>
class Book { public: int price(); };
int price(Book* b) {
  if (b == NULL) throw "NPE!";
  return (*b).price() * 1.1;
}
void print(Book* b) {
  try {
    std::cout << "The price: " << price(b);
  } catch (char const* e) {
    std::cout << "Error: " << e;
  }
}
\end{ffcode}

This mechanism may be implemented in \eolang{}:

\begin{ffcode}
[] > book
  [] > price ?
[b] > price
  if. > @
    b.eq 0
    error "NPE!"
    b.price.times 1.1
[b] > print
  try > @
    [] >>
      io.stdout > @
        tt.sprintf *1
          "The price: %d"
          price b
    [e] >>
      io.stdout > @
        tt.sprintf *1
          "Error: %s"
          e
    true
\end{ffcode}

Here, the object \ff{try} expects three arguments: 1)~an abstract object to be dataized, 2)~an abstract object to be copied with one argument, in case dataization returns an encapsulated object, and 3)~an object to be dataized anyway (similar to Java \ff{finally} block).

In the object \ff{price}, we get the object \ff{error}, which, if dataized, causes the termination of dataization and a non-conditional jump to the ``catch'' object of the \ff{try}.
The mechanism is similar to ``checked'' exceptions in Java, where a method's signature must declare all types of exceptions the method may throw.

\subsection{Many Exception Types}

\broken

A Java method may throw a number of either checked or unchecked exceptions, for example:

\begin{ffcode}
void f(int x) throws IOException {
  if (x == 0) {
    throw new IOException();
  }
  throw new RuntimeException();
}
\end{ffcode}

This would be represented in \eolang{} as such:

\begin{ffcode}
[x] > f
  if. > @
    x.eq 0
    error "IOException"
    error "RuntimeException"
\end{ffcode}

To catch both exceptions the object \ff{f} would be used like this:

\begin{ffcode}
try
  []
    try > @
      []
        f 5 > @
      [e1]
        io.stdout e1 > @
      []
        nop > @
  [e2]
    io.stdout e2 > @
  []
    nop > @
\end{ffcode}

\section{Anonymous Functions}
\label{sec:blocks}

Anonymous functions can be passed into methods as arguments. For example, in this Ruby code a ``block'' (a name Ruby uses for anonymous functions) is passed:

\begin{ffcode}
def scan(lines)
  lines.each do |t|
    if t.starts_with? '#' yield t
  end
end
scan(array) { |x| puts x }
\end{ffcode}

This mechanism may be implemented in \eolang{}:

\begin{ffcode}
[lines b] > scan
  lines.each > @
    [t] >>
      if. > @
        t.starts-with "#"
        b t
        true
scan
  array
  [x] >>
    io.stdout x > @
\end{ffcode}

Here, the anonymous function is passed to the object \ff{scan} as an argument \ff{b}. The ``call'' of this function is the dataization of its copy, performed by the \ff{if} atom.

\section{Generators}
\label{sec:generators}

\broken

A generator is a routine that can be used to control the iteration behavior of a loop.
For example, this PHP program will print the first ten Fibonacci numbers:

\begin{ffcode}
function fibonacci(int $limit):generator {
  yield $a = $b = $i = 1;
  while (++$i < $limit) {
    $b = $a + $b;
    yield $b - $a;
    $a = $b;
  }
}
foreach (fibonacci(10) as $n) {
  echo "$n\n";
}
\end{ffcode}

This mechanism may be implemented in \eolang{}:

\begin{ffcode}
[limit f] > fibonacci
  memory 0 > a
  memory 0 > b
  memory 0 > i
  seq * > @
    a.write 1
    b.write 1
    f 0
    while.
      seq (i.write (i.plus 1)) (i.lt limit)
      [idx]
        seq * > @
          b.write (a.plus b)
          a.write (b.minus a)
          f (b.minus a)
    true
fibonacci > @
  10
  [n]
    io.stdout > @
      tt.sprintf "%d\n" n
\end{ffcode}

Here, the generator is turned into an abstract object \ff{fibonacci} with an extra parameter \ff{f}, which is specified with an abstract object argument that prints what is given as \ff{n}.

\section{Types and Type Casting}
\label{sec:types}

\broken

A type system is a logical system comprising a set of rules that assign a property called a type to various constructs of a computer program, such as variables, expressions, functions, or modules.
The main purpose of a type system is to reduce possibilities for bugs in computer programs by defining interfaces between different parts of a computer program, and then checking that the parts have been connected in a consistent way.
In this example, a Java method \ff{add} expects an argument of type \ff{Book} and compilation would fail if another type is provided:

\begin{ffcode}
class Cart {
  private int total;
  void add(Book b) {
    this.total += b.price();
  }
}
\end{ffcode}

The restrictions enforced by the Java type system at compile time through types \ff{Cart} and \ff{Book} may be represented in \eolang{} using decorators, for example:

\begin{ffcode}
[] > original-cart
  memory 0 > total
  [b] > add
    total.write (total.plus (b.price))
[] > cart
  original-cart > @
  [b] > add
    if.
      b.subtype-of (tt.text "Book")
      @.add b
      []
        error "Type mismatch, Book expected" > @
\end{ffcode}

Here, it is expected that the parameter \ff{b} is defined in a ``class'' that has a \ff{subtype-of} attribute, which may also be provided by a decorator (a simplified example):

\begin{ffcode}
[] > original-book
  memory 0 > price
[] > book
  original-book > @
  [t] > subtype-of
    t.eq "Book" > @
  [] > price
    @.price > @
\end{ffcode}

This decoration may be simplified through a ``fluent'' supplementary object:

\begin{ffcode}
type "Cart" > cart-type
.super-types "Object" "Printable"
.method "add" "Book"
cart-type original-cart > cart
\end{ffcode}

Here, the \ff{type} object implements all necessary restrictions that the Java type system may provide for the type \ff{Cart} and its methods.

Type casting, which is a mechanism for changing an expression from one data type to another, may also be implemented through the same decorators.

\section{Reflection}
\label{sec:reflection}

\broken

Reflection is the ability of a process to examine, introspect, and modify its own structure and behavior.
In the following Python example, the method \ff{hello} of an instance of class \ff{Foo} is not called directly on the object but is first retrieved through reflection functions \ff{globals} and \ff{getattr}, and then executed:

\begin{ffcode}
def Foo():
    def hello(self, name):
        print("Hello, %s!" % name)
obj = globals()["Foo"]()
getattr(obj, "hello")("Jeff")
\end{ffcode}

It may be implemented in \eolang{} by encapsulating additional meta information in classes explained in \cref{sec:classes}.

\subsection{Monkey Patching}

\broken

Monkey patching is making changes to a module or class while the program is running.
This JavaScript program adds a new method \ff{print} to the object \ff{b} after the object has already been instantiated:

\begin{ffcode}
function Book(t) { this.title = t; }
var b = new Book("Object Thinking");
b.print = function() {
  console.log(this.title);
}
b.print();
\end{ffcode}

This program may be translated to \eolang{}, assuming that \ff{b} is being held by an attribute of a larger object, for example \ff{app}:

\begin{ffcode}
[] > app
  cage 0 > b
  b' > copy
  seq * > @
    b.write
      [] > book
        [t] > new
          memory "" > title
          title.write t > @
    copy.<
    b.write
      [] > book
        [t] > new
          seq * > @
            []
              copy > @
              [] > print
                io.stdout (^.title)
    b.print
\end{ffcode}

Here, the modification to the object \ff{book} happens by making a copy of it, creating a decorator, and then storing it where the original object was located.

\section{Static Methods}
\label{sec:static}

A static method (or static function) is a method defined as a member of an object but accessible directly from an API object's constructor, rather than from an object instance created via the constructor. For example, this C\# class consists of a single static method:

\begin{ffcode}
class Utils {
  public static int max(int a, int b) {
    if (a > b) return a;
    return b;
  }
}
\end{ffcode}

It may be converted to the following \eolang{} code, since a static method is nothing more than a ``global'' function:

\begin{ffcode}
[a b] > utils-max
  if. > @
    a.gt b
    a
    b
\end{ffcode}

\section{Inheritance}
\label{sec:inheritance}

\broken

Inheritance is the mechanism of basing a class upon another class while retaining similar implementation. In this Java code class \ff{Book} inherits all methods and non-private attributes from class \ff{Item}:

\begin{ffcode}
class Item {
  private int p;
  int price() { return p; }
}
class Book extends Item {
  int tax() { return price() * 0.1; }
}
\end{ffcode}

It may be represented in \eolang{} like this:

\begin{ffcode}
[] > item
  memory 0 > p
  [] > price
    p > @
[] > book
  item > i
  [] > tax
    (ms.number (i.price)).as-float.times 0.1 > @
\end{ffcode}

Here, composition is used instead of inheritance.

\subsection{Prototype-Based Inheritance}

\broken

So-called prototype-based programming uses generalized objects, which can then be cloned and extended. For example, this JavaScript program defines two objects, where \ff{Item} is the parent object and \ff{Book} is the child that inherits the parent through its prototype:

\begin{ffcode}
function Item(p) { this.price = p; }
function Book(p) {
  Item.call(this, p);
  this.tax = function () {
    return this.price * 0.1;
  }
}
var t = new Book(42).tax();
console.log(t); // prints "4.2"
\end{ffcode}

This mechanism of prototype-based inheritance may be translated to the following \eolang{} code, using the mechanism of decoration:

\begin{ffcode}
[p] > item
  memory 0 > price
  [] > new
    price.write p > @
[] > book
  [p] > new
    item p > @
    [] > tax
      times. > @
        as-float.
          ms.number (^.price)
        0.1
io.stdout
  tt.sprintf
    "%f"
    tax.
      book.new 42
\end{ffcode}

\subsection{Multiple Inheritance}

\broken

Multiple inheritance is a feature of some object-oriented computer programming languages in which an object or class can inherit features from more than one parent object or parent class. In this C++ example,
the class \ff{Jack} has both \ff{bark} and \ff{listen} methods, inherited from \ff{Dog} and \ff{Friend} respectively:

\begin{ffcode}
#include <iostream>
class Dog {
  virtual void bark() {
    std::cout << "Bark!";
  }
};
class Friend {
  virtual void listen();
};
class Jack: Dog, Friend {
  void listen() override {
    Dog::bark();
    std::cout << "Listen!";
  }
};
\end{ffcode}

It may be represented in \eolang{} like this:

\begin{ffcode}
[] > dog
  [] > bark
    io.stdout "Bark!" > @
[] > friend
  [] > listen
[] > jack
  dog > d
  friend > f
  [] > listen
    seq * > @
      d.bark
      io.stdout "listen!"
\end{ffcode}

Here, inherited methods are explicitly listed as attributes in the object \ff{jack}. This is very close to what would happen in the virtual table of a class \ff{Jack} in C++. The \eolang{} object \ff{jack} just makes it explicit.

\section{Method Overloading}
\label{sec:overloading}

\broken

Method overloading is the ability to create multiple functions with the same name but different implementations. Calls to an overloaded function will run a specific implementation of that function appropriate to the context of the call, allowing one function call to perform different tasks depending on context. In this Kotlin program two functions are defined with the same name, while only one of them is called with an integer argument:

\begin{ffcode}
fun foo(a: Int) {}
fun foo(a: Double) {}
foo(42)
\end{ffcode}

It may be represented in \eolang{} like this:

\begin{ffcode}
[args...] > foo
  (args.at 0) > a0
  if.
    a0.subtype-of "Int"
    first-foo a0
    second-foo a0
foo 42
\end{ffcode}

This code expects arguments of \ff{foo} to be equipped with the type system suggested in \cref{sec:types}. The attribute \ff{subtype-of} will help dispatch the call to the right objects.

\section{Java Generics}
\label{sec:generics}

\broken

Generics extend Java's type system to allow a type or method to operate on objects of various types while providing compile-time type safety.
For example, this Java class expects another class to be specified as \ff{T} before use:

\begin{ffcode}
class Cart<T extends Item> {
  private int total;
  void add(T i) {
    total += i.price();
  }
}
\end{ffcode}

It may be represented in \eolang{} like this:

\begin{ffcode}
[] > cart
  memory 0 > total
  [i] > add
    total.write > @
      total.plus (i.price)
\end{ffcode}

As the example demonstrates, the presence of generics in class declarations may be ignored, since \eolang{} is a language without types and type checking.

\section{C++ Templates}
\label{sec:templates}

\broken

Templates are a feature of the C++ programming language that allows functions and classes to operate with generic types, allowing a function or class to work with many different data types without being rewritten for each one.

\begin{ffcode}
template<typename T> T max(T a, T b) {
  return a > b ? a : b;
}
int x = max(7, 42);
\end{ffcode}

It may be represented in \eolang{} like this:

\begin{ffcode}
[a b] > max
  if. > @
    a.gt b
    a
    b
max 7 42 > x
\end{ffcode}

As the example demonstrates, the presence of templates may be ignored since \eolang{} is a language without types and type checking.

\section{Mixins}
\label{sec:mixins}

\broken

A mixin is a class that contains methods for use by other classes without having to be the parent class of those other classes. The following code demonstrates how a Ruby module is included in a class:

\begin{ffcode}
module Timing
  def recent?
    @time - Time.now < 24 * 60 * 60
  end
end
def News
  include Timing
  def initialize(t)
    @time = t
  end
end
n = News.new(Time.now)
n.recent?
\end{ffcode}

This code may be represented in \eolang{} by just copying the method \ff{recent?} to the object \ff{News} as if it was defined there.

\section{Annotations}
\label{sec:annotations}

\broken

In Java, an annotation is a form of syntactic metadata that can be added to source code; classes, methods, variables, parameters, and Java packages may be annotated. Later, annotations may be retrieved through the Reflection API. For example, this program analyzes the annotation attached to the class of the provided object and changes the behavior depending on one of its attributes:

\begin{ffcode}
interface Item {}
@Ship(true) class Book { /*..*/ }
@Ship(false) class Song { /*..*/ }
class Cart {
  void add(Item i) {
    /* add the item to the cart */
    if (i.getClass()
      .getAnnotation(Ship.class)
      .value()) {
      /* mark the cart as shippable */
    }
  }
}
\end{ffcode}

The code may be represented in \eolang{} using classes suggested in \cref{sec:classes}:

\begin{ffcode}
[] > book
  ship true > a1
  [] > new
    [] > @
[] > song
  ship FALSE > a1
  [] > new
    [] > @
[] > cart
  [i] > add
    # Add the item to the cart
    if. > @
      i.a1.value
        # Mark the cart shippable
        true
        FALSE
\end{ffcode}

Annotations become attributes of objects, which represent classes in \eolang{}, as explained in \cref{sec:classes}: \ff{book.a1} and \ff{song.a1}.

\section{Traceability}

\broken

A certain amount of semantic information may be lost during the translation from a more powerful programming language to \eolang{} objects, such as, for example, names, line numbers, comments, etc. Moreover, it may be useful to have the ability to trace \eolang{} objects back to the original language constructs from which they were derived. For example, a simple C function:

\begin{ffcode}
int f(int x) {
  return 42 / x;
}
\end{ffcode}

It may be mapped to the following \eolang{} object:

\begin{ffcode}
[x] > f
  [] > @
    "src/main.c:1-1" > source
    42.div x > @
  "src/main.c:0-2" > source
\end{ffcode}

Here, the synthetic \ff{source} attribute represents the location in the source code file with the original C code. It is important to ensure during translation that the name \ff{source} does not conflict with a possibly similar name in the object.

\section{Conclusion}

We demonstrated how some language features often present in high-level object-oriented languages, such as Java or C++, may be expressed using objects.
We used \eolang{} as a destination programming language because of its minimalistic semantics: it has no language features aside from objects and their composition and decoration mechanisms.

\raggedright
\bibliographystyle{ACM-Reference-Format}
\bibliography{bibliography/main}

\end{document}
